---
title: "daily-09-10"
author: "Caitlin Rasbid"
date: "2025-03-02"
format: html
execute:
  echo: true
   
---
# Question 1
```{r}
data(airquality)
?airquality
```

# Question 2
```{r}
library(visdat)
vis_dat(airquality)
vis_miss(airquality)
vis_guess(airquality)
```
## After looking at the data, I would say it needs cleaning because 5% of observations were missing, 24% of which were from Ozone and 5% of which were from Solar Radiation. Because removing 5% of observations from the datset should have minimal impacts on the model, I choose to remove n/a values.


## Cleaning the data
```{r}
library(tidyr)
library(dplyr)
library(flextable)

cleaned_airquality <- airquality %>%
  drop_na(Ozone, Solar.R)

vis_dat(cleaned_airquality)


```

# Question 3
```{r}
model <- lm(Ozone ~ Temp, data = cleaned_airquality)


```
## I initally chose to use Solar Radiation as my predictor variable. I chose this variable because Ozone concentrations in the atmosphere are correlated with the amount of solar radiation. Because of this established relationship I thought that the resulting model would be valid. The R^2 variable indicated that only 12.13% of the variance in Ozone was exlained by Solar Radiation. After using Temperature instead, another value correlated with Ozone concentration, the R^2 value was much higher and the p-values much lower.

# Question 4
```{r}
summary(model)
shapiro.test(model$residuals)
```
## I think this is a valid model because the residuals are all lower than the Solar.R model, the p-values are lower, and the R-squared value is higher.

# Question 5
## The R^2 variable is represented by two different summary stastics, the Multiple R-Squared and the Adjusted R-Squared. The multipe r-squared is the proportion of variance in the dependent variable explained by the model. In this case, it is 0.488, meaning that 48.8% of the variance in the Ozone can be explained by Temperature. The Adjusted R-squared adjusts for multiple predictors, which is why the two R-Squared values are very similar since there is only one predictor variable. 

# Question 6
```{r}
library(broom)
augmented_airquality <- augment(model)
head(augmented_airquality)
```

# Question 7
```{r}
library(ggplot2)
ggplot(augmented_airquality, aes(x = Ozone, y = .fitted)) +
  geom_point(color = "blue") +  
  geom_abline(intercept = 0, slope = 1, color = "red") +  
  labs(
    title = "Actual vs Predicted Ozone",
    subtitle = paste("Correlation:", round(cor(augmented_airquality$Ozone, augmented_airquality$.fitted), 2)),
    x = "Actual Ozone",
    y = "Predicted Ozone"
  ) +
  theme_minimal()
```

